### 2014-11-06 Publications/2013-TMI/02-abstract.tex
one more update to the paper

We propose a surface-to-volume variational solution to the registration problem,
  that exploits the detailed anatomy extracted manually or automatically
  from structural images.
We use an ``active contours without edges'' model to search for a
  deformation field of B-Spline functions that optimally maps the
  surface prior onto the data in the target space.
The strategy implicitly segments with sub-pixel resolution the image.
We demonstrate the usefulness of the approach on a digital phantom, and on
  10 real datasets randomly selected from the Human Connectome Project.

>>>>

We propose a surface-to-volume variational solution to the registration problem,
  that exploits the detailed anatomy extracted manually or automatically
  from structural images.
We use an ``active contours without edges'' model to search for a
  deformation field of B-Spline functions that optimally maps the
  surface prior onto the data in the target space.
The strategy implicitly segments with sub-pixel resolution the image.
We illustrate the usefulness of the approach on a digital phantom.
Finally, we apply the registration to correct for suscetibility distortions
  10 real diffusion MRI datasets.
The joint segmentation-registration method successfully estimated the
  distortions.
We also compare its performance to a widely-used correction method that
  demands for an extra MR acquisition.

### 2014-11-06 Publications/2013-TMI/03-introduction.tex
one more update to the paper

An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
  and reference images with an affine transformation supporting the coordinates mapping
  (the so-called ``2 \glspl*{pde} approach'').
\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
  based on the level-sets framework and using only one \gls*{pde}.
The principal application of these techniques was the image registration between timesteps in
  a time-series or between different images \citep{paragios_level_2003}.
\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
  using a free deformation field model.
A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
  functional in joint registration and segmentation, thorugh the propagation of the deformation
  ield from the contours to the whole image definition.
Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
Even though the historical evolution of joint segmentation and registration procedures is strongly
  linked to the level sets approximation, our work is based in derivations of the \gls{acwe}
  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.

>>>>

An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
  and reference images with an affine transformation supporting the coordinates mapping
  (the so-called ``2 \glspl*{pde} approach'').
\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
  based on the level-sets framework and using only one \gls*{pde}.
The principal application of these techniques was the image registration between timesteps in
  a time-series or between different images \citep{paragios_level_2003}.
\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
  using a free deformation field model.
A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
  functional in joint registration and segmentation, through the propagation of the deformation
  field from the contours to the whole image definition.
Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
Even though the historical evolution of joint segmentation and registration procedures is strongly
  linked to the level sets approximation, our work is based in derivations of the \acrlong{acwe}
  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.

### 2014-11-06 Publications/2013-TMI/04-methods.tex
one more update to the paper

Finally, the \gls{map} problem is adapted into a variational one applying the
  following log-transform:
  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
  &= C + \underset{l}{\sum}
  \underset{\vec{r} \in \Omega_l}{\sum}
  \mdist{f'}{l} \notag\\
  &+ \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2},
  \label{eq:energy}
  \end{align}
  that is the dual expression to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with anisotropic regularization as studied in
  \citep{nagel_investigation_1986}.
We provide a proof of duality in {\color{red} [REF figshare?]}.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figures/figure01.pdf}
	\caption{The active contours are defined as the interfacing surfaces of the segmenting
	\glspl{roi} $\Omega_i$, and represented in green and dark blue colors in this
	close-up.
	They iteratively through their normals $\hat{n}_i$ at each vertex $\vec{v}_i$ of the mesh.
	The gradient speeds $\vec{s}_i$ are computed as the disparity between data energies of
	the vertex $\vec{v}_i$ in each limiting region as described in \autoref{eq:energy}.
	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
	box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer.
	Finally, every $\bar{s}_i$ is projected into the grid of control points $\vec{u}_k$ that
	support the deformation field through the corresponding weights $\psi_k(\vec{v}_i)$ obtained
	as in \autoref{eq:basis_derivative}.
	}\label{fig:method}
\end{figure}

@@

\subsubsection{Optimization}
\label{sec:gradient_descent}
To find the minimum of the energy functional \eqref{eq:energy},
  we propose a gradient-descent approach with respect to the underlying
  deformation field through the following \gls*{pde}:
  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}
  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:
  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  C + \underset{l}{\sum}
  \int_{\Omega_l} \mdist{f'}{l} \notag\\
  &+ \int_{\Omega_l} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \int_{\Omega_l} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2}
  \Big\}.
  \label{eq:gradient_descent}
  \end{align}

Therefore, we can apply a discretized interpretation of \eqref{eq:shape_gradients}
  to compute the data term in \eqref{eq:gradient_descent} as follows:
  \begin{align}
  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
  \underset{l}{\sum} \frac{ \partial }{\partial \vec{u}_k} \left\{
   \underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \right\}
  = \underset{l}{\sum} \underset{i}{\sum}
   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
  \label{eq:gradient_wshape}
  \end{align}
  in this case, the formulation has been adapted to the non-binary case, $l,m$
  being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
  them such that $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
  and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.

Finally, we can compute:
  \begin{align}
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
  \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\}
  = \psi_k(\vec{v}_i)\, \hat{\vec{e}}
  \label{eq:basis_derivative}
  \end{align}
  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \autoref{eq:shape_gradients},
	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$
	corresponding to $\vec{u}_k$:
  \begin{equation}
  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{equation}
  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
  \begin{align}
  \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
  &\underset{l,m}{\sum} \underset{i}{\sum}
  \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
  -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
  \label{eq:gradient_final}
  \end{align}

\subsubsection{Region descriptor reestimation}
In regular intervals, after certain number of iterations,
the parameters $\Theta_l$ of the regions can be reestimated
based on the shifted volumetric samples
$\vec{x}' = \vec{x_0} + u(\vec{x})$.

\subsubsection{Convergence}
{\color{red} discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.}

\subsubsection{Efficient field densification}
As long as the dense deformation field is iteratively interpolated
from the same set of control points that define the $L-1$ contours,
it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
sparse matrix for fast densification. As well, we achieve a
diffeomorphic transform by

@@

\subsubsection{Simulated digital phantom} %
\label{sec:digital_phantoms}
%
We first present the proof of the concept on a simplistic digital phantom,
  simulating a spherical crust of thin tissue with a folding resembling
  a gyrus of the brain cortex folding (see \autoref{fig1}).
We simulated 2.0mm isotropic \gls*{t1} and \gls*{t2} images at
  several rician noise levels, with parameters TE/TR=10/1500ms and
  TE/TR=90/5000ms respectively using \cite{caruyer_phantomas_2014}.
The reference surfaces were extracted using the marching-cubes algorithm
  on high-resolution (1.0mm isotropic) binary versions of the corresponding
  maps of tissue volume fractions used during the simulation step.

\subsubsection{Real \gls{mri} datasets} %
\label{sec:human_connectome}
%
For the evaluation of the algorithm with real brains with \gls*{dmri} data,
  we randomly selected 10 datasets from the minimally preprocessed
  \citep{glasser_minimal_2013} cohort of the \gls*{hcp}.
These datasets comprehend the preprocessed (corrected for artifacts,
  brain-extracted and spatially normalized) \gls*{t1}, \gls*{t2} and
  multi-shell \gls*{dmri} images, along with the most prominent results of
  the standard processing with \emph{FreeSurfer}: \emph{aseg} and
  \emph{aparc} segmentations, and the surfaces corresponding to the cortical
  \gls*{gm} exterior (\emph{pial}) and the \gls*{wm}/\gls*{gm} interface
  (\emph{white}).
This cohort is especially convenient for their careful preprocessing,
  making available high-standard \gls*{mri} data accurately corrected
  for distortions and artifacts.
For the exact details of the scanning parameters we refer the reader to
  \citep{essen_human_2012}.
Selecting the appropriate labels in the \emph{aseg} segmentation, we applied
  the marching-cubes algorithm again to extract the surfaces containing the
  \gls*{csf} of the ventricular system.

As we propose the \gls*{fa} and the \gls*{md} as target features, we
  implemented a simplified pipeline for processing \gls*{dmri} using
  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
\todo[inline]{insert here the summary of parameters}

\subsection{Experiments and evaluation}
\label{sec:experiments_evaluation}
%
\subsubsection{Synthetic distortion}
\label{sec:distortion}

\subsubsection{Evaluation indices} %
\label{sec:evaluation}
We report the following indices:
  \begin{itemize}
  \item \gls{se}, a distance between the one-to-one corresponding
    vertices, weighted by their respective Voronoi area.
    \begin{equation}
    MSE = \sum\limits_k \sum\limits_j^M w_j\,\| \mathbf{x}_j - \hat{\mathbf{x}}_j \|
    \end{equation}
    where $\mathbf{x}_j$ are the locations of the $M$ vertices of the $k$ priors,
    $\hat{\mathbf{x}}_j$ are the corresponding locations recovered, and $w_j$ the
    weighting factor as the relative surface of the Voronoi area.
  \item \gls{wi}, L2-distance between the theoretical and the recovered
    deformation field.
    \begin{equation}
    WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
    \end{equation}
    where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
    and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
  \item Parcellation agreement, the Dice overlap index of the \glspl*{roi}
    contained by the parcellation.
  \end{itemize}

>>>>

Finally, the \gls{map} problem is adapted into a variational one applying the
  following log-transform:
  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
  &= C + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}
  that is the dual expression to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with anisotropic regularization as studied in
  \citep{nagel_investigation_1986}.
We provide a proof of duality in {\color{red} [REF figshare?]}.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figures/figure01}
	\caption{The active contours are defined as the interfacing surfaces of the segmenting
	\glspl{roi} $\Omega_i$, and represented in green and dark blue colors in this
	close-up.
	They iteratively through their normals $\hat{n}_i$ at each vertex $\vec{v}_i$ of the mesh.
	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
	the vertex $\vec{v}_i$ in each limiting region as described in \autoref{eq:energy}.
	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
	box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer.
	Finally, every $\bar{s}_i$ is projected into the grid of control points $\vec{u}_k$ that
	support the deformation field through the corresponding weights $\psi_k(\vec{v}_i)$ obtained
	as in \autoref{eq:basis_derivative}.
	}\label{fig:method}
\end{figure}

@@

\subsubsection{Optimization}
\label{sec:gradient_descent}
To find the minimum of the energy functional \eqref{eq:energy},
  we propose a gradient-descent approach with respect to the underlying
  deformation field through the following \gls*{pde}:
  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}
  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:
  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} ] \,d\vec{r}
  \Big\}.
  \label{eq:gradient_descent}
  \end{align}

Therefore, we can apply a discretized interpretation of \eqref{eq:shape_gradients}
  to compute the data term in \eqref{eq:gradient_descent} as follows:
  \begin{align}
  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \left\{
   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
  = \underset{i}{\sum}
   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
  \label{eq:gradient_wshape}
  \end{align}
  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
  being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
  them such that $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
  and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.

Finally, we can compute:
  \begin{align}
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
  \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\}
  = \psi_k(\vec{v}_i)\, \hat{\vec{e}}
  \label{eq:basis_derivative}
  \end{align}
  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \autoref{eq:shape_gradients},
	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$
	corresponding to $\vec{u}_k$:
  \begin{equation}
  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{equation}
  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
  \begin{align}
  \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
  &\underset{i}{\sum} \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
  -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
  \label{eq:gradient_final}
  \end{align}

\subsection{Experiments and evaluation}
\label{sec:experiments_evaluation}
%
The experiments follow a five-steps design:
\begin{enumerate}
	\item The prior surfaces are extracted from an undistorted and high-resolution dataset
	  (e.g. \autoref{fig:phantom}).
	\item A plausible, synthetic distortion $U^{-1}_{true}$ is generated.
	\item The low-resolution (target) images are warped using the synthetic distortion.
	\item Our method is run, obtaining a $\hat{U}_{tst}$.
	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
	\item Results are visually and quantitatively evaluated.
\end{enumerate}

To quantitatively evaluate the registration results we computed the \gls*{swindex} as the one-to-one
  distance between corresponding vertices of surfaces, weighted by their respective
  Voronoi area $w_i$:
  \begin{equation}
  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} w_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|
  \end{equation}
  where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
    prior, $A_p$ the total area of surface $p$, and $\hat{\vec{v}}_i$ is the location
    recovered corresponding to the vertex $\vec{v}_i$.

In the application on real data, three experts blindly rated correction results obtained
  with the proposed method and the cross-comparison reference.
The full report of visual results is provided within the Supplemental Materials.

  % \item \gls{wi}, L2-distance between the theoretical and the recovered
  %   deformation field.
  %   \begin{equation}
  %   WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
  %   \end{equation}
  %   where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
  %   and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
  % \item Parcellation agreement, the Dice overlap index of the \glspl*{roi}
  %   contained by the parcellation.
  % \end{itemize}

@@

\subsubsection{Simulated digital phantom} %
\label{sec:digital_phantoms}
%
We first present the proof of the concept on a simplistic digital phantom
  (\autoref{fig:phantom}).
Using \cite{caruyer_phantomas_2014}, we simulated 2.0mm isotropic \gls*{t1} and
  \gls*{t2} images at several rician noise levels, with parameters TE/TR=10/1500ms
  and TE/TR=90/5000ms respectively.
The reference surfaces were extracted using the marching-cubes algorithm
  shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
  on high-resolution (1.0mm isotropic) and binary images corresponding to
  the tissue fraction maps used in simulations.

\begin{figure}
	\includegraphics[width=\linewidth]{figures/figure02}
	\caption{The ``cortex'' phantom is a spherical shape with two sulci and an
	 outer crust resembling the cortical folding.
	When the model is used to generate low resolution images, the
	  partial volume effect turns segmentation of the sulci a
	  challenging problem with voxel-wise clustering methods,
	  but it is successfully segmented with the proposed method (see
	  \autoref{fig:results_phantom}).
	Top row shows the central-axial plane of the simulated images at
	  high resolution and noiseless (left is \gls*{t1} and right is
	  \gls*{t2}).
	Bottom row represents the extracted contours that are used as shape
	  priors to be registered to low-resolution, noise corrupted versions
	  of the phantom.
	}\label{fig:phantom}
\end{figure}

\subsubsection{Real \gls{mri} datasets} %
\label{sec:human_connectome}
%
For the evaluation of the algorithm to real \gls*{dmri} data of human brains,
  we downloaded 10 datasets from the ``minimally preprocessed''
	 database of the \gls*{hcp}.
We refer the reader to \citep{essen_human_2012} for exact details about acquisition
  parameters, and \citep{glasser_minimal_2013} for preprocessing issues.
The datasets comprehend \gls*{t1}, \gls*{t2} and multi-shell \gls*{dmri} images
  already corrected for artifacts, brain-extracted and spatially normalized,
  along with the most prominent results of the standard processing with
  \emph{FreeSurfer}: \emph{aparc} segmentations, and the surfaces corresponding to
  the cortical \gls*{gm} exterior and the \gls*{wm} interface with
  cortical \gls*{gm}.

### 2014-11-06 Publications/2013-TMI/06-results.tex
one more update to the paper

\section{Results and discussion}
\label{sec:results}

Please! Write me ASAP.

>>>>

% -*- root: 00-main.tex -*-
\section{Results and discussion}
\label{sec:results}

\subsection{Proof of concept}
\label{sec:results_phantom}

Applying the evaluation protocol introduced in \autoref{sec:experiments_evaluation} on
  the ``cortex'' phantom (\autoref{fig:phantom}), we successfully tested the method and
  fine tuned the algorithm for its use on real data.
In this characterization experiment, we generated random smooth displacement fields $U^{-1}_{true}$
  (step 4 in \autoref{sec:experiments_evaluation}) based on a B-Spline field with
  control points evenly located in a $6 \times 6 \times 6$ grid covering the region of the phantom.
Invertibility is ensured by controlling the maximum	displacement
  \citep{rueckert_diffeomorphic_2006}.
The software tool to generate these fields is included with the released package.
In this experiment we used $\hat{U}_{cc} = (U^{-1}_{true})^{-1}$, i.e. the ground-truth
  displacement field.
We performed the experiment in two different realizations of the phantom, one at full resolution
  and a second with half resolution to study the performance under strong partial volume effect
  conditions.
The distorted phantoms and visual results are presented in \autoref{fig:results_phantom}.
Quantitative evaluation of the \gls*{swindex} is reported in \autoref{tab:swi_phantom}.

\begin{table}
	\caption{Quantitative evaluation on phantoms: \acrlong*{swindex}}
	\label{tab:swi_phantom}
    \begin{tabular}{llll}
    Voxel size  & Interior surface & Exterior surface & Average \\
    $1.0mm$ & $0.718mm$            & $0.724mm$            & $0.721mm$   \\
    $2.0mm$ & $0.723mm$            & $0.726mm$            & $0.725mm$   \\
    \end{tabular}
\end{table}

{\color{red}
\subsubsection{Convergence evindence and settings}
 discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
\subsubsection{Efficient field densification}
As long as the dense deformation field is iteratively interpolated
from the same set of control points that define the $L-1$ contours,
it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
sparse matrix for fast densification. As well, we achieve a
diffeomorphic transform by}

% In regular intervals, after certain number of iterations,
% the parameters $\Theta_l$ of the regions can be reestimated
% based on the shifted volumetric samples
% $\vec{x}' = \vec{x_0} + u(\vec{x})$.

\begin{figure*}
\includegraphics[width=\textwidth]{figures/figure03}
\caption{Proof of concept of the registration method. First row shows the registration experiment
  on the high-resolution ($1.0mm$ isometric) realization of the phantom, second row contains the
  low-resolution ($2.0mm$ isometric) version.
  Both phantoms were simulated with \gls*{snr} 30.
  In the first column, the distorted phantoms presented with the undistorted contours.
  For the second column, the contours have been replaced by the distorted contours using
    the ground-truth warping $U^{-1}_{true}$.
  The third column contains again the contours warped with the ground-truth field
  (thin yellow line), and the corresponding contours (green and blue thick lines)
  obtained with the presented method.
  Finally, last row represents the recovered images unwarped using the $\hat{U}_{tst}$
    found through our registration.
  }\label{fig:results_phantom}
\end{figure*}

\subsection{Susceptibility distortion correction}
\label{sec:results_hcp}

As we propose the \gls*{fa} and the \gls*{md} as target features, we
  implemented a simplified pipeline for processing \gls*{dmri} using
  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
\todo[inline]{insert here the summary of parameters}

Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
  the marching-cubes algorithm again to extract priors of the following
  homogeneous regions in terms of \gls*{fa} and \gls*{md} joint distribution:
\begin{itemize}
	\item \gls*{csf} of the ventricular system,
	\item deep \gls*{gm} structures,
	\item \gls*{wm} surface as the ensemble of brain stem, and
	  cerebellar and cerebral \gls*{wm},
	\item cortical \gls*{gm} surface, including cerebellar \gls*{gm}.
\end{itemize}

\subsection{Discussion}
\label{sec:discussion}
We propose a registration-segmentation algorithm designed for biomedical
  image analyses that are affected by all or some of the following circumstances:
  \begin{itemize}
  	\item Reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably, or there is not available an apt image
 		to be used as reference.
  	\item The low resolution of the target volume produces strong partial volume effects
  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
  	registration.
  	\item The unknown warping is smooth enough to be represented with a sufficient number
  	of BSpline kernels, and its explicit need is justified (i.e. the distortion
  	field in \gls*{dmri} images, object tracking applications, etc.).
  \end{itemize}
The algorithm is designed for the proposed application on correcting distortions of
  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
  networks.
As a consequence, our approach relies on the availability of precise segmentations or
  surfaces of the objects that are to be segmented on the target volume(s).
Connectome extraction protocols generally include the acquisition of \gls*{t1} images
  to provide with prior information about anatomy with high accuracy.
Our method solves in a single step a joint problem usually addressed in two steps.
The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data (i.e.
  \cite{greve_accurate_2009}), whereas the second corrects for the nonlinear distortions
  derived from the inhomogenety of magnetic susceptibility across tissues
  \citep{jezzard_correction_1995}.

We implemented the registration method upon the widely used Insight Registration and Segmentation
	Toolkit (\url{http://www.itk.org}) with the aim to release a useful and free software bundle.
The instrumentation framework is also distributed with the registration method,
  and has been implemented using \emph{nipype} \citep{gorgolewski_nipype_2011}.
The pipelines include the automatic generation of the ``cortex'' phantom.
The real datasets are publicly available under the \gls*{hcp} project.
Thus, all the experiments and results presented in this paper are intended to be
  fully reproducible.
